The problem and promise of scale in multilayer animal social networks

### 2020-02-15 paper/manuscript.Rmd
typos

## Landscape data and habitat classification
Landcover data were provided by the Newfoundland and Labrador Wildlife Division
[@Integrated_2013]. Available landcover classification included nine habitat
types at 30 m resolution. We reclassified the landcover types into three
categories: foraging habitat, open habitat, and forest habitat using the R
package `raster` [@Hijmans_2019]. Foraging habitat consisted of lichen habitat,
open habitat consisted of wetland, rocky barrens, and anthropogenic habitat
types, while forest habitat consisted of coniferous forest, conifer scrub,
broadleaf forest, and mixed-wood forest. Water habitat was excluded from all
subsequent analyses.

@@

### Network types and construction
We generated a multilayer social network consisting `r DT[, uniqueN(get(idcol))]`
individual nodes and two aspects (season and landcover). Two season
elementary-layers (summer and winter) and three landcover elementary-layers
(forage, forest, and open) combined for a total of six layers. Single layers were
composed of individuals lnked by social associations, for each season and landcover combination. Proximity-based social
network layers were generated using the R package `spatsoc` [@Robitaille_2019]
in `r version$version.string` [@R_Core_Team_2019].

@@

(ref:ml) Visualization of multilayer network for social ungulate case study representing social association of caribou (*Rangifer tarandus*) on Fogo Island, Newfoundland in summer 2017 and winter 2018. Individual caribou (n = `r DT[, uniqueN(get(idcol))]` females) are show as nodes and each indiviual is represented by a unique color. Six layers represent combinations of two seasons (summer and winter) and three habitat classes (open, forest and forage). Intralayer edges connect pairs of individuals that associated according to spatial and temporal thresholds described above, and are scaled in line thickness according to strength of association (SRI). Interlayer edges connect individuals to themselves across seasons and within habitat classes (e.g. individual "A" in summer, forest and individual "A" in winter, forest).

@@

Some multilayer network metrics were strongly influenced by the length of the
time window. Mean graph strength was weak but became slightly stronger with
increasing time window length during the summer, but the higher mean graph
strength during the winter showed a slight increase then decrease with
increasing winter length (Figure \@ref(fig:winlen)a). Layer relevance increased
for all summer layers as individuals became more connected with increasing
window lengths, but winter layers showed relatively stable but high layer
relevance (Figure \@ref(fig:winlen)b). This pattern indicates that the
proportion of neighbours that were the same across layers increased with a
longer time frame during the summer, butthere were a high proportion of
neighbors that remained the same across layers during the winter, regardless of
time length. Furthermore, though the layers were generally not very similar
between the summer and winter, the length of the time window created slight
variability in this metric, particularly for forage and open habitat types
(Figure \@ref(fig:winlen)c). Connective redundancy increased marginally with
increasing window length, indicating that the proportions of neighbours that
were the same across layers, though relatively high in general, slightly
increased when time length was greater (Figure \@ref(fig:winlen)d). These
results suggest that an increasing window length, especially for the summer
season, may be resulting in less precisely defined seasons that are not clearly
capturing periods of aggregation and dissaggregation related to resource
availability.

@@

<!-- Figure: metrics by landcover resolution -->
(ref:lcres) Varying spatial scale in landcover resolution. For each landcover resolution, multilayer networks were constructed of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland in summer 2017 and winter 2018 across three habitat classes (forage, forest, and open). A) Mean graph strength of individuals in summer layers had consistent weak connections. Individuals in winter forest and open habitats increased consistently while winter forage varied with a peak at 250 m and valley at 500 m. B) Mean layer relevance showed sparse connectivity in summer compared to winter networks and a decrease in winter forage layer relevance at higher landcover resolutions. C) Similarity of landcover network layers between seasons (e.g. open summer compared to open winter) orage and forest habitats but showed more variation in open habitats with increasing social thresholds. Note that layer similarity in forage is not shown because individuals did not interact at landcover resolutions greater than 30 m during the summer in forage habitats, so layer similarity could not be calculated. D) Mean connective redundancy across individuals showed a slight decreasing trend with increasing landcover resolution.

@@

The integration of behavioural and ecological theory has shown the value of
considering the importance of various ecological disciplines to better
understand and predict animal social networks, including demography
[@Shizuka_2019], evolutionary and quantitative genetics [@Fisher_2017], disease
ecology [@Silk_2017], and animal personality [@Sih_2018]. While the number of
studies integrating social network theory into a broader array of behavioural
and ecological sub-disciplines is impressive, only recently has the role of
animal space use been considered in describing the causes and consequences of
the social environment in the context of (1) phenotypes and individual fitness
[@Webber_2018], (2) movement ecology and collective movement [@Bode_2011], and
(3) habitat selection and space use [@He_2019]. Importantly, much of the work
linking space use and the social environment has been theoretical and empirical
studies are lacking [but see @Peignier_2019; @Strandburg_Peshkin_2015;
@Strandburg_Peshkin_2017]. Here, we discuss the role of scale and mutlilayer
networks in the emerging integration of social and spatial ecology within the
behavioural ecology literature.

@@

Conversely, picking the proper temporal scale is important but a challenge
because all layers in the multilayer network should reflect the same time
period, and the period should be long enough for each layer to have enough data
for robust networks [@Farine_2017]. These considerations may be limiting factors
when answering questions that require timeframes that are relevant to different
individuals, for instance. In both mono- and multilayer networks, the effect of
scale on the relationship between an individual’s social network position and
fitness need to be considered based on the hypotheses being tested and
logistical constraints.

@@

In contrast to spatial networks, where nodes represent spatial locations that
are fixed in space, movement networks can represent continuous and simultaneous
movement trajectories of multiple animals. In a recent study of vulturine
gunieafowl (*Acryllium vulturinum*), GPS movement data were used to generate
monolayer association networks within a multilayer network [@Papageorgiou_2019].
However, despite potential to generate networks in different spatial or social
contexts, layers in the multilayer network represented social associations
within and between groups. Continuous collection of animal relocation data using
GPS technology can be used to measure how groups of animals make collective
movement decisions in the context of their local environment
[@Strandburg_Peshkin_2015; @Strandburg_Peshkin_2017] and across scales. A
collective movement layer could constitute similarity in movement for dyads
[@Long_2014] or groups [@Bode_2011]. This movement layer could be made at a
fine [i.e. minute-to-minute: @Cleasby_2019] or coarse scale [global migration:
@Flack_2018]. Our case study integrates variation in temporal and spatial scale
to assess the role of scale in predicting multilayer networks. As one might
expect, we found increasing social threshold results in saturation of potential
social connections (Figure \@ref(fig:socres)).<!--TODO: what??, while increasing temporal scale had little effect
on habitat-specific networks (Figure \@ref(fig:))-->. Another potential application would be
to generate social networks during times when animals are engaged in different
behaviours, including traveling, interacting, or foraging [@Muller_2018;
@Finn_2019] and construct a multilayer network based on behaviour-specific
monolayer networks [e.g. @Smith_Aguilar_2019].

@@

Given the existing framework for constructing multilayer networks from
behavioural layers [e.g. @Smith_Aguilar_2019], monolayer networks that explicity
quantify movement paramters could be itnegrated. For example, edges could
represent measures of dyadic or collective movement based on dynamic interaction
metrics [@Long_2014]. Integrating movement layers with habitat-specific social
association layers could further illuminate the role of movement as a driver
of the social and spatial environments an individual experiences [@Webber_2018].
In our case study, we generated multilayer networks based on habitat-specific
monolayers. Coordinated movement of dyads, or groups, could vary based on
habitat, and the approach we outline could be used to explore mechanisms linking
the social and spatial environments. For example, social processes may be an
emergent property of the landscape because animals aggregate at resources, such
as waterholes [@Chamaill_Jammes_2008]. Movement to, from, or within habitat
patches or territories can also contribute to the formation of the social
environment [@Spiegel_2016]. For example, in sleepy lizards (*Tiliqua rugosa*)
males interacted primarily at home range edges, presumably driving territorial
behaviour [@Spiegel_2018]. Incorporating movement ecology within a multilayer
network framework is a logical next step for a burgeoning field. In particular,
when association networks are habitat-specific, we suggest the integration of
movement and space use in multilayer networks provides novel insights into the
effects of habitat configuration on the formation of the social environment
[@He_2019; @Webber_2018].

@@

The emergent relationship between habitat and sociality has been explored by the
distribution and phenology of resources and use of social information. Cougar
(*Puma concolor*) home ranges overlapped more in regions with high hunting
opportunity, a proxy for resource richness [@Elbroch_2015]. Caribou (*Rangifer
tarandus*) sociality differed between seasons, where social associations were
stronger in winter and contrastingly low in summer when resource availability
are more homogenous [@Peignier_2019]. Bison (*Bison bison*) showed stronger
habitat selection in larger groups and avoidance of predation risk was a main
factor contributing to their habitat selection [@Fortin_2009]. These examples,
comparing sociality and habitat selection across temporal or spatial contexts,
are implicitly multilayer analyses.

<!-- TODO: weak -->
Multilayer networks can be used to explicitly consider habitat selection and
sociality across scales. Animal social systems can be parsed by spatial
contexts, defined, for example, by habitat types as in Figure \@ref(fig:lcres).
The effect of seasonality of resources can can be directly measured by differences in
habitat selection and sociality across temporal layers. Habitat selection can be
explicitly integrated using network layers defined by similarity of selection
coefficients between individuals. These explicit integrations of habitat
selection and sociality across social, spatial and temporal scales highlight the
potential for novel insights from multilayer networks.

@@

Silk et al. (2018) summarized some key multilayer questions in animal behaviour
research. While these questions are important for linking animal behaviour and
multilayer networks, we summarize our perspectives on the future of multilayer
networks with an additional set of questions posed through the lens of spatial
ecology that are relevant to the promise and problem of scale in mutlilayer
networks:

>>>>

## Landscape data and habitat classification
Land cover data were provided by the Newfoundland and Labrador Wildlife Division
[@Integrated_2013]. Available landcover classification included nine habitat
types at 30 m resolution. We reclassified the landcover types into three
categories: foraging habitat, open habitat, and forest habitat using the R
package `raster` [@Hijmans_2019]. Foraging habitat consisted of lichen habitat,
open habitat consisted of wetland, rocky barrens, and anthropogenic habitat
types, while forest habitat consisted of coniferous forest, conifer scrub,
broad leaf forest, and mixed-wood forest. Water habitat was excluded from all
subsequent analyses.

@@

### Network types and construction
We generated a multilayer social network consisting `r DT[, uniqueN(get(idcol))]`
individual nodes and two aspects (season and landcover). Two season
elementary-layers (summer and winter) and three landcover elementary-layers
(forage, forest, and open) combined for a total of six layers. Single layers were
composed of individuals linked by social associations, for each season and landcover combination. Proximity-based social
network layers were generated using the R package `spatsoc` [@Robitaille_2019]
in `r version$version.string` [@R_Core_Team_2019].

@@

(ref:ml) Visualization of multilayer network for social ungulate case study representing social association of caribou (*Rangifer tarandus*) on Fogo Island, Newfoundland in summer 2017 and winter 2018. Individual caribou (n = `r DT[, uniqueN(get(idcol))]` females) are show as nodes and each individual is represented by a unique color. Six layers represent combinations of two seasons (summer and winter) and three habitat classes (open, forest and forage). Intralayer edges connect pairs of individuals that associated according to spatial and temporal thresholds described above, and are scaled in line thickness according to strength of association (SRI). Interlayer edges connect individuals to themselves across seasons and within habitat classes (e.g. individual "A" in summer, forest and individual "A" in winter, forest).

@@

Some multilayer network metrics were strongly influenced by the length of the
time window. Mean graph strength was weak but became slightly stronger with
increasing time window length during the summer, but the higher mean graph
strength during the winter showed a slight increase then decrease with
increasing winter length (Figure \@ref(fig:winlen)a). Layer relevance increased
for all summer layers as individuals became more connected with increasing
window lengths, but winter layers showed relatively stable but high layer
relevance (Figure \@ref(fig:winlen)b). This pattern indicates that the
proportion of neighbours that were the same across layers increased with a
longer time frame during the summer, but there were a high proportion of
neighbors that remained the same across layers during the winter, regardless of
time length. Furthermore, though the layers were generally not very similar
between the summer and winter, the length of the time window created slight
variability in this metric, particularly for forage and open habitat types
(Figure \@ref(fig:winlen)c). Connective redundancy increased marginally with
increasing window length, indicating that the proportions of neighbours that
were the same across layers, though relatively high in general, slightly
increased when time length was greater (Figure \@ref(fig:winlen)d). These
results suggest that an increasing window length, especially for the summer
season, may be resulting in less precisely defined seasons that are not clearly
capturing periods of aggregation and dissaggregation related to resource
availability.

@@

<!-- Figure: metrics by landcover resolution -->
(ref:lcres) Varying spatial scale in landcover resolution. For each landcover resolution, multilayer networks were constructed of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland in summer 2017 and winter 2018 across three habitat classes (forage, forest, and open). A) Mean graph strength of individuals in summer layers had consistent weak connections. Individuals in winter forest and open habitats increased consistently while winter forage varied with a peak at 250 m and valley at 500 m. B) Mean layer relevance showed sparse connectivity in summer compared to winter networks and a decrease in winter forage layer relevance at higher landcover resolutions. C) Similarity of landcover network layers between seasons (e.g. open summer compared to open winter) forage and forest habitats but showed more variation in open habitats with increasing social thresholds. Note that layer similarity in forage is not shown because individuals did not interact at landcover resolutions greater than 30 m during the summer in forage habitats, so layer similarity could not be calculated. D) Mean connective redundancy across individuals showed a slight decreasing trend with increasing landcover resolution.

@@

The integration of behavioural and ecological theory has shown the value of
considering the importance of various ecological disciplines to better
understand and predict animal social networks, including demography
[@Shizuka_2019], evolutionary and quantitative genetics [@Fisher_2017], disease
ecology [@Silk_2017], and animal personality [@Sih_2018]. While the number of
studies integrating social network theory into a broader array of behavioural
and ecological sub-disciplines is impressive, only recently has the role of
animal space use been considered in describing the causes and consequences of
the social environment in the context of (1) phenotypes and individual fitness
[@Webber_2018], (2) movement ecology and collective movement [@Bode_2011], and
(3) habitat selection and space use [@He_2019]. Importantly, much of the work
linking space use and the social environment has been theoretical and empirical
studies are lacking [but see @Peignier_2019; @Strandburg_Peshkin_2015;
@Strandburg_Peshkin_2017]. Here, we discuss the role of scale and multilayer
networks in the emerging integration of social and spatial ecology within the
behavioural ecology literature.

@@

Conversely, picking the proper temporal scale is important but a challenge
because all layers in the multilayer network should reflect the same time
period, and the period should be long enough for each layer to have enough data
for robust networks [@Farine_2017]. These considerations may be limiting factors
when answering questions that require time frames that are relevant to different
individuals, for instance. In both mono- and multilayer networks, the effect of
scale on the relationship between an individual’s social network position and
fitness need to be considered based on the hypotheses being tested and
logistical constraints.

@@

In contrast to spatial networks, where nodes represent spatial locations that
are fixed in space, movement networks can represent continuous and simultaneous
movement trajectories of multiple animals. In a recent study of vulturine
gunieafowl (*Acryllium vulturinum*), GPS movement data were used to generate
monolayer association networks within a multilayer network [@Papageorgiou_2019].
However, despite potential to generate networks in different spatial or social
contexts, layers in the multilayer network represented social associations
within and between groups. Continuous collection of animal relocation data using
GPS technology can be used to measure how groups of animals make collective
movement decisions in the context of their local environment
[@Strandburg_Peshkin_2015; @Strandburg_Peshkin_2017] and across scales. A
collective movement layer could constitute similarity in movement for dyads
[@Long_2014] or groups [@Bode_2011]. This movement layer could be made at a
fine [i.e. minute-to-minute: @Cleasby_2019] or coarse scale [global migration:
@Flack_2018]. Our case study integrates variation in temporal and spatial scale
to assess the role of scale in predicting multilayer networks. As one might
expect, we found increasing social threshold results in saturation of potential
social connections (Figure \@ref(fig:socres)).<!--TODO: what??, while increasing temporal scale had little effect
on habitat-specific networks (Figure \@ref(fig:))-->. Another potential application would be
to generate social networks during times when animals are engaged in different
behaviours, including travelling, interacting, or foraging [@Muller_2018;
@Finn_2019] and construct a multilayer network based on behaviour-specific
monolayer networks [e.g. @Smith_Aguilar_2019].

@@

Given the existing framework for constructing multilayer networks from
behavioural layers [e.g. @Smith_Aguilar_2019], monolayer networks that explicitly
quantify movement parameters could be integrated. For example, edges could
represent measures of dyadic or collective movement based on dynamic interaction
metrics [@Long_2014]. Integrating movement layers with habitat-specific social
association layers could further illuminate the role of movement as a driver
of the social and spatial environments an individual experiences [@Webber_2018].
In our case study, we generated multilayer networks based on habitat-specific
monolayers. Coordinated movement of dyads, or groups, could vary based on
habitat, and the approach we outline could be used to explore mechanisms linking
the social and spatial environments. For example, social processes may be an
emergent property of the landscape because animals aggregate at resources, such
as waterholes [@Chamaill_Jammes_2008]. Movement to, from, or within habitat
patches or territories can also contribute to the formation of the social
environment [@Spiegel_2016]. For example, in sleepy lizards (*Tiliqua rugosa*)
males interacted primarily at home range edges, presumably driving territorial
behaviour [@Spiegel_2018]. Incorporating movement ecology within a multilayer
network framework is a logical next step for a burgeoning field. In particular,
when association networks are habitat-specific, we suggest the integration of
movement and space use in multilayer networks provides novel insights into the
effects of habitat configuration on the formation of the social environment
[@He_2019; @Webber_2018].

@@

The emergent relationship between habitat and sociality has been explored by the
distribution and phenology of resources and use of social information. Cougar
(*Puma concolor*) home ranges overlapped more in regions with high hunting
opportunity, a proxy for resource richness [@Elbroch_2015]. Caribou (*Rangifer
tarandus*) sociality differed between seasons, where social associations were
stronger in winter and contrastingly low in summer when resource availability
are more homogeneous [@Peignier_2019]. Bison (*Bison bison*) showed stronger
habitat selection in larger groups and avoidance of predation risk was a main
factor contributing to their habitat selection [@Fortin_2009]. These examples,
comparing sociality and habitat selection across temporal or spatial contexts,
are implicitly multilayer analyses.

<!-- TODO: weak -->
Multilayer networks can be used to explicitly consider habitat selection and
sociality across scales. Animal social systems can be parsed by spatial
contexts, defined, for example, by habitat types as in Figure \@ref(fig:lcres).
The effect of seasonality of resources can can be directly measured by differences in
habitat selection and sociality across temporal layers. Habitat selection can be
explicitly integrated using network layers defined by similarity of selection
coefficients between individuals. These explicit integration of habitat
selection and sociality across social, spatial and temporal scales highlight the
potential for novel insights from multilayer networks.

@@

Silk et al. (2018) summarized some key multilayer questions in animal behaviour
research. While these questions are important for linking animal behaviour and
multilayer networks, we summarize our perspectives on the future of multilayer
networks with an additional set of questions posed through the lens of spatial
ecology that are relevant to the promise and problem of scale in multilayer
networks:
